---
title: "Obtaining the Raw Data"
author: "Paul A. Rubin"
date: "01/06/2015"
output: html_document
---

This file documents how the data for the course
project in ["Getting and Cleaning Data"](https://class.coursera.org/getdata-010) (Coursera, section 10, January 2015) was obtained.

To execute the code here, first modify the working directory in the `setwd()` command as appropriate, then run or knit the document. Note that the R chunks are all set not to evaluate by default. To run the script (rather than just knitting it), change the global evaluation option to TRUE in the following chunk.

```{r}
# decide whether R chunks should be evaluated
library(knitr)
opts_chunk$set(eval = FALSE)
```
The date the data was obtained is shown above.

# Steps to obtain the data:

First, change to the working directory (previously created).

```{r}
setwd("~/Desktop/project")
```

Next, set the URL based on the project instructions.

```{r}
dataURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
```

Now download the file.

```{r}
download.file(dataURL, destfile = "data.zip", method = "curl")
```

Unzip the file into the 'data' subdirectory of
the working directory (creating it if necessary).

```{r}
unzip("data.zip", exdir = "data", setTimes = TRUE)
```

Confirm the contents of the data directory.

```{r}
list.files(path = "data", recursive = TRUE, all.files = TRUE)
```
